Fusions in Rewriting: Foundation of Left-linear Graph Rewriting Systems

### 2024-02-03 icalp/fusione.tex
title
switch e controesempi

\usepackage{hyperref}
\usepackage[textsize=tiny]{todonotes}
%\usepackage{etex}
\usepackage[all]{xy}
\SelectTips{cm}{}
% derivations and labels
\usepackage{proof}
\newcommand{\lab}[1]{\ensuremath{\mathsf{{#1}}}}
\newcommand{\slab}[1]{\ensuremath{\scriptstyle{\mathsf{{#1}}}}}

@@

\title{Switch equivalence for left-linear rewriting systems}
\author{Paolo Baldan} 
{Department of Mathematics, University of Padova, Italy}
{baldan@math.unipd.it}{}{}
\author{Davide Castelnovo}
{Department of Mathematics, University of Padova, Italy}
{davide.castelnovo@math.unipd.it}{}{}
\author{Andrea Corradini}
{Department of Computer Science, University of Pisa, Italy}
{andrea.corradini@unipi.it}{}{}

@@

\begin{definition}
	Let $\X$ be a category and $\mathcal{A}$ a  subclass of $\mor(\X)$. We say that  $\mathcal{A}$ is
	\begin{itemize}
		\item 
		\emph{stable under pushouts (pullbacks)} if for every pushout (pullbacks) square 
		\[\xymatrix{A\ar[r]^f  \ar[d]_{m}& B \ar[d]^n \\ C \ar[r]_g & D}\]
		if $m \in \mathcal{A}$ ($n\in \mathcal{A}$) then $n \in \mathcal{A}$ ($m \in \mathcal{A}$);
		\item \emph{closed under composition} if $g, f\in \mathcal{A}$ implies $g\circ f\in \mathcal{A}$ whenever $g$ and $f$ are composable;
		\item \emph{closed under decomposition} if $g\circ f, g\in \mathcal{A}$ implies $f\in \mathcal{A}$.
	\end{itemize}
\end{definition}

@@

The following proposition establishes a key property of $\mathcal{A}$-Van Kampen squares with a mono as a side: they are not only pushouts, but also pullbacks.
\begin{theoremEnd}[category=sec2]{proposition}\label{prop:pbpo} Let $\mathcal{A}$ be a class of arrows stable under pushouts and containing all the isomorphisms.  If $m\colon A\to C$ is mono and belongs to $\mathcal{A}$, then every $\mathcal{A}$-Van Kampen square
	\[\xymatrix{A\ar[r]^{g} \ar[d]_{m} & B \ar[d]^{n} \\ C \ar[r]_{f}  & D}\]
	is also a pullback square and $n$ is a monomorphism.
\end{theoremEnd}
\begin{proof}\begin{proofEnd}
		We can start considering the cube below and noticing that $n$, being the pushout of $m\in \mathcal{M}$, belongs to $\mathcal{A}$ too.
		\[\xymatrix@C=13pt@R=13pt{&A\ar[dd]|\hole_(.65){\id{A}}\ar[rr]^{g} \ar[dl]_{\id{A}} && B \ar[dd]^{\id{B}} \ar[dl]_{\id{B}} \\ A  \ar[dd]_{m}\ar[rr]^(.65){g} & & B \ar[dd]_(.3){n}\\&A\ar[rr]|\hole^(.65){g} \ar[dl]^{m} && B \ar[dl]^{n} \\C \ar[rr]_{f} & & D}\]
		By construction the top face of the cube is a pushout and the back one a pullback. The left face is a pullback because $m$ is mono. Thus the $\mathcal{A}$-Van Kampen property yields that the front and the right faces are pullbacks. 
	\end{proofEnd} 
\end{proof}

@@

@@

\begin{definition}[Sequential independence]
  \label{de:sequential-independence}
  Let $(\X, \R)$ be a left-linear DPO-rewriting system with $\X$ an $\mathcal{M}$-adhesive category. Let also $\dder{D}\colon G\Mapsto H$ and $\dder{D'}\colon H\Mapsto T$ be the two direct derivations depicted below.

	\[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h} & L' \ar[d]_{m'}& K' \ar[d]^{k'}\ar@{>->}[l]_{l'} \ar[r]^{r'} & R' \ar[d]^{h'}\\G & \ar@{>->}[l]^{f} D \ar[r]_{g}& H & H & \ar@{>->}[l]^{f'} D' \ar[r]_{g'}& T}\]

	An \emph{independence pair} between $\dder{D}$ and $\dder{D'}$, is a pair of  arrows $i_1\colon R\to D'$ and $i_2\colon L'\to D$ such that the following diagram commutes.

	\[\xymatrix@C=15pt{L \ar[d]_{m}&& K \ar[d]_{k}\ar@{>->}[ll]_{l} \ar[r]^{r} & R \ar@/^.35cm/@{.>}[drrr]|(.3)\hole_(.4){i_0} \ar[dr]|(.3)\hole_{h} && L' \ar@/_.35cm/@{.>}[dlll]^(.4){i_1} \ar[dl]|(.3)\hole^{m'}& K' \ar[d]^{k'}\ar@{>->}[l]_{l'} \ar[rr]^{r'} && R' \ar[d]^{h'}\\G && \ar@{>->}[ll]^{f} D \ar[rr]_{g}&& H  && \ar@{>->}[ll]^{f'} D' \ar[rr]_{g'}&& T}\]
	We will say that $\dder{D}$ and $\dder{D'}$ are \emph{weakly sequentially independent} if an independence pair exists. If such independence pair is unique we will say that $\dder{D}$ and $\dder{D'}$ are \emph{sequentially independent}.
\end{definition}

@@

 Let $(i_0, i_1)$ and $(i'_0, i'_1)$ be independence pairs for the direct derivations $\dder{D}$ and $\dder{D'}$. Notice that, by definition, we have
 \[f'\circ i_0=h=f'\circ i'_0\]
	On the other hand, 
	$f'\colon D'\to H$ is  the pushout of $l'\colon K'\to L'$ and so it is in $\mathcal{M}$, implying $i_0=i'_0$. If, moreover, we suppose that the rule $\rho$ applied in $\dder{D}$ is linear, then $g\colon D\to H$ is in $\mathcal{M}$ too, hence, from the equation
	\[
	g\circ i_1=m'= g\circ i'_1\]
	we can deduce that $i'_0=i'_1$, too. Summing up, we have just proved the following.

@@

We want to be able to exchange to switch the application of two rules used by two consecutive weakly independent direct derivations. This leads us to the following definition.

@@

 	Take another derivation $\der{E}=\{\dder{E}_i\}_{i=0}^1$ with source $G_0$ and target $G_2$ and in which $\der{E}_0$, $\der{E}_1$ are
\[\xymatrix{L_{\der{E},0} \ar[d]_{m_{\der{E},0}}& K_{\der{E},0} \ar[d]^{k_{\der{E},0}}\ar@{>->}[l]_{l_{\der{E},0}} \ar[r]^{r_{\der{E},0}} & R_{\der{E},0} \ar[d]^{h_{\der{E},0}} & L_{\der{E},1} \ar[d]_{m_{\der{E},1}}& K_{\der{E},1} \ar[d]^{k_{\der{E},1}}\ar@{>->}[l]_{l_{\der{E},1}} \ar[r]^{r_{\der{E},1}} & R_{\der{E},1} \ar[d]^{h_{\der{E},1}}\\G_{0} & \ar@{>->}[l]^{f_{\der{E},0}} D_{\der{E},0} \ar[r]_{g_{\der{E},0}}& G_{\der{E},1} & G_{\der{E},1} & \ar@{>->}[l]^{f_{\der{E},1}} D_{\der{E},1} \ar[r]_{g_{\der{E},1}}& G_{2}}\]
 	We say that $\der{E}$ is a
 	 	 \emph{switch} of $\dder{D}_0$ and $\dder{D}_1$ along $(i_1, i_2)$ if
 	\begin{enumerate}
 		\item $\dder{E}_0$ applies the rule used by $\dder{D}_1$ and $\dder{E}_1$ the one used by $\dder{D}_0$, so that
 		\[(l_{\der{E},0}, r_{\der{E},0})=(l_1,r_1)  \qquad (l_{\der{E},1}, r_{\der{E},1})=(l_0,r_0)\]
 		\item there exists an independence pair $(j_0, j_1)$ between $\dder{E}_0$ and $\dder{E}_1$ such that
 		\[ m_0=f_{\der{E},0} \circ j_1 \qquad h_1=g_{\der{E},1} \circ j_0\]
 		\item $m_{\der{E},0}= f_0\circ i_1$ and $h_{\der{E},1}= g_{1}\circ i_0$.
 	\end{enumerate}

@@

\begin{remark}\label{rem:fact}
	Notice that, since $f_{\der{E}, 0}$ is in $\mathcal{M}$, there is at most one $j_1\colon L_0\to D_{\der{E},0}$ such that $m_0=f_{\der{E}, 0} \circ j_1$. Moreover, we already know that there is at most one $j_0\colon R_1\to D_{\der{E},1}$ such that $m_{\der{E},1}=f_{\der{E},1}\circ j_0$, therefore, the independence pair $(j_0,j_1)$ between the component of a switch is uniquely determined.
\end{remark}

Consider a sqitch $\der{E}=\{\dder{E}_i\}_{i=0}^1$ for $\dder{D}_0$ and $\dder{D}_1$, then by definition $\dder{E}_0$ and $\dder{E}_1$ are weakly sequentially independentWe can then wonder if they are switchable.

\begin{theoremEnd}[category=sec3]{proposition}
	Let $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation and suppose that $\der{E}=\{\dder{E}_i\}_{i=0}^1$ is a switch for it. Then $\der{D}$ is a switch for $\dder{E_0}$ and $\dder{E}_1$.
\end{theoremEnd}
\begin{proof}\begin{proofEnd}
	The first point of \Cref{def:switc} is obviously satisfied. The second one follows at once since 
	\[m_{\der{E},0}= f_0\circ i_1 \qquad h_{\der{E},1}= g_{1}\circ i_0\]
	Finally, the third point holds because we already know that 
\[m_{\der{E},0}= f_0\circ i_1 \qquad h_{\der{E},1}= g_{1}\circ i_0\]
	\end{proofEnd}
\end{proof}

@@

\begin{lemma}[Uniqueness of switches]\label{thm:switch_uni} Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^1$ be a derivation and suppose that both $\der{E}=\{\dder{E}_i\}_{i=0}^1$  and $\der{F}=\{\dder{F}_i\}_{i=0}^1$ are switches of $\dder{D}_0$ and $\dder{D}_1$. Then $\der{E}$ and $\der{F}$ are abstraction equivalent.
\end{lemma}
\begin{proof}
	By definition of switch we know that $\der{E}_0$ and $\der{F}_0$ have the same match $f_0\circ i_1$ and that $\der{E}_1$ and $\der{F}_1$ have the same comatch $g_1\circ i_0$. Thus we can build the solid part of the diagram below.
	\[\xymatrix@C=15pt{G_{0} && \ar@{>->}[ll]_{f_{\der{F},0}} D_{\der{F},0} \ar[rr]^{g_{\der{F},0}}&& G_{\der{F},1} && \ar@{>->}[ll]_{f_{\der{F},1}} D_{\der{F},1} \ar[rr]^{g_{\der{F},1}}&& G_{2}\\L_1 \ar[d]^{f_0\circ i_1} \ar[u]_{f_0\circ i_1}&& K_{1} \ar[d]^{k_{\der{E},0}} \ar[u]_{k_{\der{F},0}}\ar@{>->}[ll]_{l_1} \ar[r]^{r_1} & R_{1} \ar@/^.35cm/[drrr]|(.3)\hole_(.4){j_0} \ar[dr]|(.3)\hole_{h_{\der{E},0}} \ar@/_.35cm/[urrr]|(.3)\hole^(.4){a_0} \ar[ur]|(.3)\hole^{h_{\der{F},0}}& &L_{0} \ar@/_.35cm/[dlll]^(.4){j_1} \ar[dl]|(.3)\hole^{m_{\der{E},1}} \ar@/^.35cm/[ulll]_(.4){a_1} \ar[ul]|(.3)\hole_{m_{\der{F},1}}& K_{0} \ar[d]_{k_{\der{E},1}} \ar[u]^{k_{\der{F},1}}\ar@{>->}[l]_{l_{0}} \ar[rr]^{r_0} && R_{0} \ar[d]_{g_1\circ i_0} \ar[u]^{g_1\circ i_0}\\ \ar@/^.5cm/[uu]^{\id{G_0}} G_{0} && \ar[ll]^{f_{\der{E},0}} D_{\der{E},0} \ar@{.>}@/^.5cm/[uu]^(.33){\phi_{0}}|\hole \ar[rr]_{g_{\der{E},0}}&& G_{\der{E},1} \ar@{.>}[uu]^(.2){\psi_1}|(.45)\hole |(.55)\hole&& \ar@{>->}[ll]^{f_{\der{E},1}} D_{\der{E},1} \ar@{.>}@/_.5cm/[uu]_(.33){\phi_{1}}|\hole\ar[rr]_{g_{\der{E},1}}&& G_{2}\ar@{.>}@/_.5cm/[uu]_{\psi_2}}\]
	Now, by \Cref{prop:unique} we get $\phi_0\colon D_{\der{E},0}\to D_{\der{F},0}$ and $\psi_1 \colon G_{\der{E},1}\to G_{\der{F},1}$ such that 
	\[\xymatrix{G_{0} & \ar@{>->}[l]_{f_{\der{F},0}} D_{\der{F},0} \ar[r]^{g_{\der{F},0}}& G_{\der{F},1} \\L_1 \ar[d]^{f_0\circ i_1} \ar[u]_{f_0\circ i_1}& K_{1} \ar[d]^{k_{\der{E},0}} \ar[u]_{k_{\der{F},0}}\ar[l]_{l_1} \ar[r]^{r_1} & R_{1}  \ar[d]_{h_{\der{E},0}}  \ar[u]^{h_{\der{F},0}}\\ \ar@/^.5cm/[uu]^{\id{G_0}} G_{0} & \ar@{>->}[l]^{f_{\der{E},0}} D_{\der{E},0} \ar@/^.5cm/[uu]^(.33){\phi_{0}}|\hole \ar[r]_{g_{\der{E},0}}& G_{\der{E},1} \ar@/_.5cm/[uu]_{\psi_1}}\]
	commutes. But then we have
	\[
	f_{\der{F},0}\circ \phi_0\circ j_1=f_{\der{E},0} \circ j_1=m_0=f_{\der{F},0}\circ a_1\]
	which entails $a_1= \phi_0\circ j_1$. This, in turn, gives us that
	\[\psi\circ m_{\der{E},1}=\psi \circ g_{\der{E},0}\circ j_1=g_{\der{F},0}\circ \phi_0\circ j_1=g_{\der{F},0}\circ a_1=m_{\der{F},1}\]

>>>>

\usepackage{hyperref}
\usepackage[textsize=tiny]{todonotes}
%\usepackage{etex}
\usepackage[all]{xy}
\SelectTips{cm}{}
% derivations and labels
\usepackage{proof}
\newcommand{\lab}[1]{\ensuremath{\mathsf{{#1}}}}
\newcommand{\slab}[1]{\ensuremath{\scriptstyle{\mathsf{{#1}}}}}
\usepackage{wrapfig}

@@

\title{Fusions in Rewriting: Foundation of Left-linear Graph Rewriting Systems}
\author{Paolo Baldan} 
{Department of Mathematics, University of Padova, Italy}
{baldan@math.unipd.it}{}{}
\author{Davide Castelnovo}
{Department of Mathematics, University of Padova, Italy}
{davide.castelnovo@math.unipd.it}{}{}
\author{Andrea Corradini}
{Department of Computer Science, University of Pisa, Italy}
{andrea.corradini@unipi.it}{}{}

@@

\begin{definition}
	Let $\X$ be a category and $\mathcal{A}$ a  subclass of $\mor(\X)$. We say that  $\mathcal{A}$ is:
	\begin{itemize}
		\parbox{11cm}{\item 
		\emph{stable under pushouts (pullbacks)} if for every pushout (pullbacks) square as the one on the right, 	if $m \in \mathcal{A}$ ($n\in \mathcal{A}$) then $n \in \mathcal{A}$ ($m \in \mathcal{A}$);
			\item \emph{closed under composition} if $g, f\in \mathcal{A}$ implies $g\circ f\in \mathcal{A}$ whenever $g$ and $f$ are composable; }\hfill
		\parbox{2cm}{$\xymatrix{A\ar[r]^f  \ar[d]_{m}& B \ar[d]^n \\ C \ar[r]_g & D}$}
		\parbox{11cm}{}\hfill

		\item \emph{closed under decomposition} if $g\circ f, g\in \mathcal{A}$ implies $f\in \mathcal{A}$.
	\end{itemize}
\end{definition}

@@

The following proposition establishes a key property of $\mathcal{A}$-Van Kampen squares with a mono as a side: they are not only pushouts, but also pullbacks.

\begin{theoremEnd}[category=sec2]{proposition}\label{prop:pbpo} Let $\mathcal{A}$ be a class of arrows stable under pushouts and containing all the isomorphisms.  If $m\colon A\to C$ is mono and belongs to $\mathcal{A}$, then every $\mathcal{A}$-Van Kampen square
		\[\xymatrix{A\ar[r]^{g} \ar[d]_{m} & B \ar[d]^{n} \\ C \ar[r]_{f}  & D}\]
	is also a pullback square and $n$ is a monomorphism.
\end{theoremEnd}

\begin{proof}\begin{proofEnd}
		We can start considering the cube below and noticing that $n$, being the pushout of $m\in \mathcal{M}$, belongs to $\mathcal{A}$ too.
		\[\xymatrix@C=13pt@R=13pt{&A\ar[dd]|\hole_(.65){\id{A}}\ar[rr]^{g} \ar[dl]_{\id{A}} && B \ar[dd]^{\id{B}} \ar[dl]_{\id{B}} \\ A  \ar[dd]_{m}\ar[rr]^(.65){g} & & B \ar[dd]_(.3){n}\\&A\ar[rr]|\hole^(.65){g} \ar[dl]^{m} && B \ar[dl]^{n} \\C \ar[rr]_{f} & & D}\]
		By construction the top face of the cube is a pushout and the back one a pullback. The left face is a pullback because $m$ is mono. Thus the $\mathcal{A}$-Van Kampen property yields that the front and the right faces are pullbacks. 
	\end{proofEnd} 
\end{proof}

@@

\begin{example}\label{rem:iso} Let $\mathsf{Iso}$ be the class of all isomorphisms, then any category $\X$ is $\mathsf{Iso}$-adhesive.
\end{example}

@@

\begin{definition}[Sequential independence]
  \label{de:sequential-independence}
  Let $(\X, \R)$ be a left-linear DPO-rewriting system with $\X$ an $\mathcal{M}$-adhesive category. Let also $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation of length $2$. 	An \emph{independence pair} between $\dder{D}_0$ and $\dder{D}_0$, is a pair of  arrows $i_0\colon R_0\to D_1$ and $i_1\colon L_1\to D_0$ such that the following diagram commutes.
	\[\xymatrix@C=15pt{L_0 \ar[d]_{m_0}&& K_0 \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0 \ar@{.>}@/^.35cm/[drrr]|(.3)\hole_(.4){i_0} \ar[dr]|(.3)\hole_{h_0} && L_1 \ar@{.>}@/_.35cm/[dlll]^(.4){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_{g_0}&& G_1  && \ar@{>->}[ll]^{f_1} D_1 \ar[rr]_{g_1}&& G_2}\]
	We will say that $\dder{D}_0$ and $\dder{D}_1$ are \emph{weakly sequentially independent} if an independence pair exists. If such independence pair is unique we will say that $\dder{D}$ and $\dder{D'}$ are \emph{sequentially independent}.
\end{definition}

@@

 Let $(i_0, i_1)$ and $(i'_0, i'_1)$ be independence pairs for the direct derivations $\dder{D}_0$ and $\dder{D}_1$. Notice that, by definition, we have
 \[f_1\circ i_0=h_0=f_1\circ i'_0\]
	On the other hand, 
	$f_1\colon D_\to G_1$ is  the pushout of $l_1\colon K_1\to L_1$ and so it is in $\mathcal{M}$, implying $i_0=i'_0$. If, moreover, we suppose that the rule $\rho$ applied in $\dder{D}$ is linear, then $g_0\colon D_0\to G_1$ is in $\mathcal{M}$ too, hence, from the equation
	\[
	g_0\circ i_1=m_1= g_0\circ i'_1\]
	we can deduce that $i'_0=i'_1$, too. Summing up, we have just proved the following.

@@

There is another useful properties of weakly sequentially independent direct derivations which is worth mentioning.

\begin{theoremEnd}[category=sec2]{proposition}\label{prop:tec}
	Let $(i_0, i_1)$ be an independence pair between two direct derivations $\dder{D}_0$ and $\dder{D}_1$ in a left-linear DPO-rewriting system $(\X,\R)$. Consider the pullback of $f_1\colon D_1\to G_1$ along $g_0\colon D_0\to G_1$ (first square below),  then there exist the arrows $u_1\colon K_1\to P$ and $u_0\colon K_0\to P$ fitting in the central and right square, Moreover, the right one is always a pullback.
	\[\xymatrix{P \ar@{>->}[r]^{p_0} \ar[d]_{p_1} & D_0 \ar[d]^{g_0}& K_0 \ar[r]^{r_0}  \ar@{.>}[d]_{u_0}& R_0 \ar[d]^{i_0} & K_1 \ar@{>->}[r]^{l_1}  \ar@{.>}[d]_{u_1}& L_1 \ar[d]^{i_1}\\  D_1 \ar@{>->}[r]_{f_1} &G_1 &P \ar[r]_{p_1} & D_1 &P \ar[r]_{p_0}  & D_0}\]
\end{theoremEnd}
\begin{proof}
	\begin{proofEnd}
We start noticing that
\[
f_1\circ i_0\circ r_0=h_0\circ r_0=g_0\circ k_0 \qquad g_0\circ i_1\circ l_1=m_1\circ l_1= f_1 \circ k_1 \] 
Thus there exists the dotted $u_0\colon K_0\to P$, $u_1\colon K_1\to P$  in the diagram
\[\xymatrix{K_0 \ar[d]_{r_0} \ar@{.>}[r]_{u_0}\ar@/^.4cm/[rr]^{k_0} &P \ar@{>->}[r]_{p_0} \ar[d]_{p_1}& D_0 \ar[d]^{g_0}&K_1 \ar@{>->}[d]_{l_1} \ar@{.>}[r]_{u_1}\ar@/^.4cm/[rr]^{k_1} &P \ar[r]_{p_1} \ar@{>->}[d]_{p_0}& D_1 \ar@{>->}[d]^{f_1}\\R_0 \ar@/_.4cm/[rr]_{h_0}\ar[r]^{i_0}& D_1 \ar@{>->}[r]^{f_1}& G_1&L_1 \ar@/_.4cm/[rr]_{m_1}\ar[r]^{i_1}& D_0 \ar[r]^{g_0}& G_1}\]
By \Cref{prop:pbpoad,lem:pb1} we get that the left half of the second rectangle is a pullback. 
	\end{proofEnd} \qedhere 
\end{proof}

We want to be able to to switch the application of two rules used by two consecutive weakly independent direct derivations. This leads us to the following definition.

@@

 	Take another derivation $\der{E}=\{\dder{E}_i\}_{i=0}^1$ as below
\[\xymatrix@C=18pt{L_{\der{E},0} \ar[d]_{m_{\der{E},0}}&& K_{\der{E},0} \ar[d]_{k_{\der{E},0}}\ar@{>->}[ll]_{l_{\der{E},0}} \ar[r]^{r_{\der{E},0}} & R_{\der{E},0} \ar[dr]_{h_{\der{E},0}} && L_{\der{E},1} \ar[dl]^{m_{\der{E},1}}& K_{\der{E},1} \ar[d]^{k_{\der{E},1}}\ar@{>->}[l]_{l_{\der{E},1}} \ar[rr]^{r_{\der{E},1}} && R_{\der{E},1} \ar[d]^{h_{\der{E},1}}\\G_{0} && \ar@{>->}[ll]^{f_{\der{E},0}} D_{\der{E},0} \ar[rr]_{g_{\der{E},0}}&& G_{\der{E},1} && \ar@{>->}[ll]^{f_{\der{E},1}} D_{\der{E},1} \ar[rr]_{g_{\der{E},1}}&& G_{2}}\]
 	We say that $\der{E}$ is a
 	 	 \emph{switch} of $\dder{D}_0$ and $\dder{D}_1$ along $(i_1, i_2)$ if
 	\begin{enumerate}
 		\item $\dder{E}_0$ applies the rule used by $\dder{D}_1$ and $\dder{E}_1$ the one used by $\dder{D}_0$, so that
 		\[(l_{\der{E},0}, r_{\der{E},0})=(l_1,r_1)  \qquad (l_{\der{E},1}, r_{\der{E},1})=(l_0,r_0)\]
 		\item there exists an independence pair $(j_0, j_1)$ between $\dder{E}_0$ and $\dder{E}_1$ such that
 		\[ m_0=f_{\der{E},0} \circ j_1 \qquad h_1=g_{\der{E},1} \circ j_0\]
 		\item $m_{\der{E},0}= f_0\circ i_1$ and $h_{\der{E},1}= g_{1}\circ i_0$.
 	\end{enumerate}

@@

\begin{remark}\label{rem:fact}
	Notice that, since $f_{\der{E}, 0}$ is in $\mathcal{M}$, there is at most one $j_1\colon L_0\to D_{\der{E},0}$ such that $m_0=f_{\der{E}, 0} \circ j_1$. Moreover, we already know that there is at most one $j_0\colon R_1\to D_{\der{E},1}$ such that $m_{\der{E},1}=f_{\der{E},1}\circ j_0$, therefore, the independence pair $(j_0,j_1)$ between the component of a switch is uniquely determined.
\end{remark}
\begin{example}\label{ex:diff1} Recall that, given two posets $(P, \leq)$ and $(Q, \leq)$, we can construct the poset $(P, \leq)\oplus(Q,\leq)$ which has as underlying set the disjoint union $P+Q$ and in which every element of $Q$ is greater then every element of $P$, while elements of $P$ and $Q$ are compared using the original orders.

	Take now the set $S:=\{a,b,c\}$ and endow it with the smallest order such that $a\leq b$ and $a\leq c$. 
	Take also the set $\mathbb{N}$ ordered by $\geq$, so that $0$ is its maximum. 
	Let $\X$ be the category associated to the order $(S, \leq)\oplus (\mathbb{N}, \geq)$ which, by \Cref{rem:iso} is $\mathsf{Iso}$-adhesive. Notice that in this category the arrows $(a,b)\colon a\to b$ $(a,c)\colon a\to c$ do not have a pushout because $b$ and $c$ do not have a supremum.  Define $\R$ as  the set containing the following two rules
	\[\xymatrix{a & a \ar[r]^{(a, 0)} \ar[l]_{\id{a}} & 0 & a & a \ar[r]^{(a, b)} \ar[l]_{\id{a}} & b  }\]
We can then consider the following derivation $\der{D}=\{\dder{D}_i\}_{i=0}^1$ of length $2$, made by two sequential independent direct derivations.
	 	\[\xymatrix@C=30pt{a \ar[d]_{m_0}&& a \ar[d]_{k_0}\ar[ll]_{\id{a}} \ar[r]^{(a,0)} & 0 \ar@/^.35cm/[drrr]|(.315)\hole_(.45){\id{0}} \ar[dr]|(.3)\hole_{\id{0}} && a \ar@/_.35cm/[dlll]^(.45){(a,c)} \ar[dl]|(.3)\hole^{(a,0)}& a \ar[d]^{(a,0)}\ar@{>->}[l]_{\id{a}} \ar[rr]^{(a,b)} && b \ar[d]^{(b,0)}\\c && \ar@{>->}[ll]^{(\id{c})} c \ar[rr]_{(c,0)}&& 0  && \ar@{>->}[ll]^{\id{0}} 0 \ar[rr]_{\id{0}}&& 0}\]

	 	On the other hand, the rule applied by $\dder{D}_1$ cannot be applied to $c$: indeed the unique morphism $a\to c$ is $(a,c)$, thus we get a diagram
	 	\[\xymatrix{a  \ar[d]_{(a,c)}& a \ar[d]_{(a,c)}\ar@{>->}[l]_{\id{a}} \ar[r]^{(a,b)} & b \\c & c \ar[l]^{\id{c}} }\]
	 	But $b$ and $c$ do not have a supremum in the poset underlying $\X$, thus the previous diagram cannot be completed to a direct derivation.
\end{example}

Consider a switch $\der{E}=\{\dder{E}_i\}_{i=0}^1$ for $\dder{D}_0$ and $\dder{D}_1$, then by definition $\dder{E}_0$ and $\dder{E}_1$ are weakly sequentially independent. We can then wonder if they are switchable.

\begin{theoremEnd}[category=sec3]{proposition}
	Let $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation and suppose that $\der{E}=\{\dder{E}_i\}_{i=0}^1$ is a switch for it. Then $\der{D}$ is a switch for $\dder{E}_0$ and $\dder{E}_1$ along $(j_0, j_1)$.
\end{theoremEnd}
\begin{proof}\begin{proofEnd}
	The first point of \Cref{def:switc} is obviously satisfied. The second one follows at once since 
	\[m_{\der{E},0}= f_0\circ i_1 \qquad h_{\der{E},1}= g_{1}\circ i_0\]
	Finally, the third point holds because we already know that 
\[m_{\der{E},0}= f_0\circ i_1 \qquad h_{\der{E},1}= g_{1}\circ i_0\]
	\end{proofEnd}
\end{proof}

@@

\begin{theoremEnd}[category=sec3]{lemma}[Uniqueness of switches]\label{thm:switch_uni} Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^1$ be a derivation and suppose that both $\der{E}=\{\dder{E}_i\}_{i=0}^1$  and $\der{F}=\{\dder{F}_i\}_{i=0}^1$ are switches of $\dder{D}_0$ and $\dder{D}_1$. Then $\der{E}$ and $\der{F}$ are abstraction equivalent.
\end{theoremEnd}
\begin{proof}
	\begin{proofEnd}		
		By definition of switch we know that $\der{E}_0$ and $\der{F}_0$ have the same match $f_0\circ i_1$ and that $\der{E}_1$ and $\der{F}_1$ have the same comatch $g_1\circ i_0$. Thus we can build the solid part of the diagram below.
		\[\xymatrix@C=22pt{G_{0} && \ar@{>->}[ll]_{f_{\der{F},0}} D_{\der{F},0} \ar[rr]^{g_{\der{F},0}}&& G_{\der{F},1} && \ar@{>->}[ll]_{f_{\der{F},1}} D_{\der{F},1} \ar[rr]^{g_{\der{F},1}}&& G_{2}\\L_1 \ar[d]^{f_0\circ i_1} \ar[u]_{f_0\circ i_1}&& K_{1} \ar[d]^{k_{\der{E},0}} \ar[u]_{k_{\der{F},0}}\ar@{>->}[ll]_{l_1} \ar[r]^{r_1} & R_{1} \ar@/^.35cm/[drrr]|(.3)\hole_(.4){j_0} \ar[dr]|(.3)\hole_{h_{\der{E},0}} \ar@/_.35cm/[urrr]|(.3)\hole^(.4){a_0} \ar[ur]|(.3)\hole^{h_{\der{F},0}}& &L_{0} \ar@/_.35cm/[dlll]^(.4){j_1} \ar[dl]|(.3)\hole^{m_{\der{E},1}} \ar@/^.35cm/[ulll]_(.4){a_1} \ar[ul]|(.3)\hole_{m_{\der{F},1}}& K_{0} \ar[d]_{k_{\der{E},1}} \ar[u]^{k_{\der{F},1}}\ar@{>->}[l]_{l_{0}} \ar[rr]^{r_0} && R_{0} \ar[d]_{g_1\circ i_0} \ar[u]^{g_1\circ i_0}\\ \ar@/^.5cm/[uu]^{\id{G_0}} G_{0} && \ar[ll]^{f_{\der{E},0}} D_{\der{E},0} \ar@{.>}@/^.5cm/[uu]^(.33){\phi_{0}}|\hole \ar[rr]_{g_{\der{E},0}}&& G_{\der{E},1} \ar@{.>}[uu]^(.76){\psi_1}|(.45)\hole |(.55)\hole&& \ar@{>->}[ll]^{f_{\der{E},1}} D_{\der{E},1} \ar@{.>}@/_.5cm/[uu]_(.33){\phi_{1}}|\hole\ar[rr]_{g_{\der{E},1}}&& G_{2}\ar@{.>}@/_.5cm/[uu]_{\psi_2}}\]
		Now, by \Cref{prop:unique} we get $\phi_0\colon D_{\der{E},0}\to D_{\der{F},0}$ and $\psi_1 \colon G_{\der{E},1}\to G_{\der{F},1}$ such that 
		\[\xymatrix{G_{0} & \ar@{>->}[l]_{f_{\der{F},0}} D_{\der{F},0} \ar[r]^{g_{\der{F},0}}& G_{\der{F},1} \\L_1 \ar[d]^{f_0\circ i_1} \ar[u]_{f_0\circ i_1}& K_{1} \ar[d]^{k_{\der{E},0}} \ar[u]_{k_{\der{F},0}}\ar[l]_{l_1} \ar[r]^{r_1} & R_{1}  \ar[d]_{h_{\der{E},0}}  \ar[u]^{h_{\der{F},0}}\\ \ar@/^.5cm/[uu]^{\id{G_0}} G_{0} & \ar@{>->}[l]^{f_{\der{E},0}} D_{\der{E},0} \ar@/^.5cm/[uu]^(.33){\phi_{0}}|\hole \ar[r]_{g_{\der{E},0}}& G_{\der{E},1} \ar@/_.5cm/[uu]_{\psi_1}}\]
		commutes. But then we have
		\[
		f_{\der{F},0}\circ \phi_0\circ j_1=f_{\der{E},0} \circ j_1=m_0=f_{\der{F},0}\circ a_1\]
		which entails $a_1= \phi_0\circ j_1$. This, in turn, gives us that
		\[\psi\circ m_{\der{E},1}=\psi \circ g_{\der{E},0}\circ j_1=g_{\der{F},0}\circ \phi_0\circ j_1=g_{\der{F},0}\circ a_1=m_{\der{F},1}\]
